---
title: "Reproducible Scientific Reporting in R and LaTeX"
output:
  html_document:
    toc: true
    toc_depth: 2
---

# Overview

This workflow demonstrates how to generate reproducible analytical outputs from a structured modeling dataset. The analysis loads simulated longitudinal data, fits statistical models, creates publication-style figures, and exports LaTeX-ready tables for integration into a scientific report.

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)

set.seed(123)
```

## Packages

```{r packages}
if (!require("pacman")) install.packages("pacman")

pacman::p_load(
  tidyverse,
  data.table,
  here,
  glmmTMB,
  modelsummary,
  broom.mixed,
  ggeffects,
  kableExtra
)
```

# Load Data

```{r load-data}
df_models <- fread(
  here("data", "simulated_df_models.csv")
)

glimpse(df_models)
```

# Prepare Data

```{r prepare-data}
df_models <- df_models %>%
  mutate(
    id = factor(id),
    group = factor(group),
    period = factor(period)
  )
```


# Exploratory Visualization

```{r exploratory-figure, fig.width=7, fig.height=5}
p_behavior <- df_models %>%
  ggplot(aes(rank_prop, behavior_count)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(
    title = "Behavior Counts by Dominance Rank",
    x = "Proportional Dominance Rank",
    y = "Behavior Count"
  ) +
  theme_minimal()

p_behavior
```

```{r save-exploratory-figure}
dir.create(here("figures"), recursive = TRUE, showWarnings = FALSE)

ggsave(
  filename = here("figures", "behavior_by_rank.png"),
  plot = p_behavior,
  width = 7,
  height = 5
)
```


# Statistical Modeling

```{r fit-models}
m_rank <- glmmTMB(
  behavior_count ~ rank_z + offset(log(time_observed)) + (1 | id),
  family = nbinom2,
  data = df_models
)

m_rank_color <- glmmTMB(
  behavior_count ~ rank_z + rg_z + offset(log(time_observed)) + (1 | id),
  family = nbinom2,
  data = df_models
)
```


# Model Summary Table

```{r model-summary-table}
dir.create(here("tables"), recursive = TRUE, showWarnings = FALSE)

models <- list(
  "Rank only" = m_rank,
  "Rank + RG" = m_rank_color
)

coef_map <- c(
  "rank_z" = "Dominance rank (z)",
  "rg_z" = "RG opponency (z)"
)

modelsummary(
  models,
  coef_map = coef_map,
  statistic = "({std.error})",
  stars = TRUE,
  output = here("tables", "model_summary.tex")
)
```


# Predicted Effects

```{r predicted-effects, fig.width=6, fig.height=5}
pred_rg <- ggpredict(
  m_rank_color,
  terms = "rg_z"
)

p_pred_rg <- ggplot() +
  geom_point(
    data = df_models,
    aes(
      x = rg_z,
      y = behavior_count,
      color = rg_z
    ),
    alpha = 0.8,
    size = 2.2
  ) +
  geom_ribbon(
    data = pred_rg,
    aes(
      x = x,
      ymin = conf.low,
      ymax = conf.high
    ),
    alpha = 0.25,
    fill = "#B22222"
  ) +
  geom_line(
    data = pred_rg,
    aes(x = x, y = predicted),
    linewidth = 1,
    color = "#7A0000"
  ) +
  scale_color_gradient(
    low = "#FFD6D6",
    high = "#7A0000"
  ) +
  labs(
    title = "Predicted Behavioral Counts",
    x = "Standardized RG Opponency",
    y = "Behavior Count",
    color = "RG Opponency"
  ) +
  theme_minimal()

p_pred_rg
```

```{r save-predicted-effects}
ggsave(
  filename = here("figures", "predicted_rg_effect.png"),
  plot = p_pred_rg,
  width = 6,
  height = 5
)
```


# Export Complete Results

```{r export-results}
dir.create(here("output"), recursive = TRUE, showWarnings = FALSE)

fwrite(
  broom.mixed::tidy(m_rank_color),
  here("output", "model_coefficients.csv")
)
```

# Summary

This workflow illustrates an end-to-end reproducible reporting pipeline: simulated longitudinal data are analyzed using mixed-effects models, results are exported as LaTeX-ready tables, and figures are saved as publication-ready files.